\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-07-17 papers/methods/paper.tex
Updating from some of Dans feedback

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{Steve}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\newcommand{\green}{\color{green}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

@@

\begin{table}
	\centering
	\caption{Top-level parameters defined in our model and a summary of their use. The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}
	\label{tab:param_summary}
	\begin{tabular}{ll}
		\hline
		Parameter & Description \\
		\hline
		$\Omega_m$  & Matter density  \\
		$w$  &  Dark energy equation of state  \\
		$\alpha$ & Stretch standardisation   \\
		$\beta$         &  Colour standardisation   \\
		$\delta(0)$ & Scale of the mass-magnitude correction\\
		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
		$\delta\mathcal{Z}_i$ & Systematics scale\\
		$\delta S$ & Selection effect deviation \\
		$\langle M_B \rangle$ & Mean absolute magnitude \\
		$s$ & Colour bias\\
		\hline
		$\langle x_1^i \rangle$ & Mean stretch nodes\\
		$\langle c^i \rangle$ & Mean colour nodes\\
		$\sigma_{M_B}$ & Population magnitude scatter \\
		$\sigma_{x_1}$ & Population stretch scatter \\
		$\sigma_{c}$ & Population colour scatter \\
		$\kappa_{0}$ & Extra colour dispersion\\
		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics. The quoted value for $w$ represents the average mean of the fits, with the uncertainty being the average standard deviation of the fits. In brackets, the simulation scatter (the standard deviation of the mean of 100 fits) is shown, and the bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation. With systematics enabled, both the {\gten} and {\celeven} models show evidence of bias, scattering to either side of the simulated value of $w=-1$. However, their deviation from the truth value represents a shift of approximately $0.3\sigma$ when taking into account the uncertainty on fits to $w$. The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations. We also note that the simulations do not vary cosmological parameters nor population, and as our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and in fact represents a minimum bound on permissible uncertainty values.}
	\label{tab:bulk_summary}
	\begin{tabular}{l|cc}
		\hline
		Model & $w$ (scatter) & Bias Significance \\
		\hline
		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$ & $3.9\sigma$  \\
		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$ & $4.4\sigma$  \\
		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$ & $2.0\sigma$   \\
		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$ & $7.9\sigma$   \\
		\hline
	\end{tabular}
\end{table}

@@

Cosmological bias is detected with a confidence of more than $3\sigma$ for the {\gten} and {\celeven} models. We investigate the bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics. This is confirmed by replacing the fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model. We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$. Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.

@@

For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}. For our full systematics model, the bias represents a deviation from $-0.15\sigma$ to $0.27\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.

>>>>

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{Steve}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}
\newcommand{\green}{\color{forestgreen}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

@@

\begin{table}
	\centering
	\caption{{\green Top-level parameters defined in our model and a summary of their use. The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}}
	\label{tab:param_summary}
	\begin{tabular}{ll}
		\hline
		Parameter & Description \\
		\hline
		$\Omega_m$  & Matter density  \\
		$w$  &  Dark energy equation of state  \\
		$\alpha$ & Stretch standardisation   \\
		$\beta$         &  Colour standardisation   \\
		$\delta(0)$ & Scale of the mass-magnitude correction\\
		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
		$\delta\mathcal{Z}_i$ & Systematics scale\\
		$\delta S$ & Selection effect deviation \\
		$\langle M_B \rangle$ & Mean absolute magnitude \\
		$s$ & Colour bias\\
		\hline
		$\langle x_1^i \rangle$ & Mean stretch nodes\\
		$\langle c^i \rangle$ & Mean colour nodes\\
		$\sigma_{M_B}$ & Population magnitude scatter \\
		$\sigma_{x_1}$ & Population stretch scatter \\
		$\sigma_{c}$ & Population colour scatter \\
		$\kappa_{0}$ & Extra colour dispersion\\
		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics. The quoted value for $w$ represents the average mean of the fits, with the uncertainty being the average standard deviation of the fits. In brackets, the simulation scatter (the standard deviation of the mean of 100 fits) is shown, and the bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation. With systematics enabled, both the {\gten} and {\celeven} models show evidence of bias, scattering to either side of the simulated value of $w=-1$. However, their deviation from the truth value represents a shift of approximately $0.3\sigma$ when taking into account the uncertainty on fits to $w$. {\green The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations. We also note that the simulations do not vary cosmological parameters nor population, and as our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and in fact represents a minimum bound on permissible uncertainty values.}}
	\label{tab:bulk_summary}
	\begin{tabular}{l|cc}
		\hline
		Model & $w$ (scatter) & Bias Significance \\
		\hline
		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$ & $3.9\sigma$  \\
		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$ & $4.4\sigma$  \\
		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$ & $2.0\sigma$   \\
		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$ & $7.9\sigma$   \\
		\hline
	\end{tabular}
\end{table}

@@

Cosmological bias is detected with a confidence of more than $3\sigma$ for the {\gten} and {\celeven} models. We investigate the bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics. This is confirmed by replacing the fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased. {\green Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model. We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$. Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.}

@@

For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}. {\green For our full systematics model, the bias represents a deviation between $-0.3\sigma$ to $0.3\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
